### 2017-08-29 sec-model.tex
Minor wording update for clarity in model section.

An HMM is a specific probabilistic generative model with a ``built-in''
state transition system that would control the data to be generated by the
model, thus it is especially suitable for modeling sequence
data~\cite{Rabiner:1990:RSR, Huang:1990:HMM}. At any moment, the HMM would
be in one of $k$ states $U=\{u_1,\ldots,u_k\}$, and at the next moment, the
HMM would move to another state stochastically according to a transition
matrix that specifies the probability $p(u_i \mid u_j)$ of going to state
$u_i$ when the HMM is currently in state $u_j$. When the HMM is in state
$u$, the HMM can generate an observable data point $x$ according to an
output probabilistic model $p(x \mid u)$. Thus, if we ``run'' an HMM for $N$
time points denoted by $t=1,\ldots,N$, the HMM could ``generate'' a sequence
of observations $x_1 \ldots x_N$, where each $x_i$ is an output symbol by
going through a sequence of \emph{hidden states} $w_1 \ldots w_N$ where
$w_i \in U$ is a state. The association of such a latent sequence of state
transitions with the observed symbols makes it possible to use the HMM to
``decode'' the latent behavior of students behind the surface behavior we
directly observe in the log data, allowing for understanding student
behavior more deeply than a model with no latent state variables.

>>>>

An HMM is a specific probabilistic generative model with a ``built-in''
state transition system that would control the data to be generated by the
model, thus it is especially suitable for modeling sequence
data~\cite{Rabiner:1990:RSR, Huang:1990:HMM}. At any moment, the HMM would
be in one of $k$ states $U=\{u_1,\ldots,u_k\}$, and at the next moment, the
HMM would move to another state stochastically according to a transition
matrix that specifies the probability $p(u_i \mid u_j)$ of going to state
$u_i$ when the HMM is currently in state $u_j$. When the HMM is in state
$u$, the HMM can generate an observable data point $x$ according to an
output probabilistic model $p(x \mid u)$. Thus, if we ``run'' an HMM for $N$
time points denoted by $t=1,\ldots,N$, the HMM could ``generate'' a sequence
of observations $x_1 \ldots x_N$, where each $x_i$ is an output symbol by
going through a sequence of \emph{hidden states} $w_1 \ldots w_N$ where
each $w_i$ is a random variable taking a value from the HMM's state set
$U$. The association of such a latent sequence of state transitions with
the observed symbols makes it possible to use the HMM to ``decode'' the
latent behavior of students behind the surface behavior we directly observe
in the log data, allowing for understanding student behavior more deeply
than a model with no latent state variables.
